Temporal ordering and registration of images in studies of developmental dynamics

### 2014-05-29 image_analysis_paper/pnas_draft2.tex
Added edits to paper

Imaging studies provide unique insights into dynamics of developmental processes. 
%
A large number of imaging studies are based on a cross-sectional experimental design, where developmental progress is arrested in a population of embryos, each of which is at a different point along its developmental trajectory. 
%
The goal is then to temporally order the data to reconstruct a representative developmental dynamics from snapshots provided by the collection of embryos.
%
Images of different biological samples must first be registered before they can be temporally ordered.
%
When such data sets are large, noisy, and/or if the developmental changes are subtle, these tasks can be difficult to perform manually.
%
We present an automated approach to registering and temporally ordering data from such cross-sectional studies.
%
The approach is based on vector diffusion maps, a broadly applicable manifold learning technique that does not require {\it a priori} knowledge of image features or developmental dynamics.
%
%Furthermore, these techniques can separate multiple trajectories within the same data set.
%
We demonstrate the utility of these methods using a collection of images from a cross-sectional study of cell signaling during dorsoventral patterning of the {\it Drosophila} embryo.
\end{abstract}

@@

Image registration has been studied in a variety of contexts \cite{zitova2003image, rowley1998rotation, hajnal2010medical, greenspan1994rotation, zhao2003face}.
%
We use a manifold learning algorithm \cite{singer2012vector} which simultaneously addresses the problems of registration and temporal ordering. 
%
This algorithm is one of many nonlinear dimensionality reduction techniques that have been developed over the past decade \cite{Belkin2003, coifman2005geometric, coifman2006geometric, tenenbaum2000global, roweis2000nonlinear}. 
%
This technique has been used for data mining in
applications ranging from analysis of cryo-electron microscopy images of individual molecules to face recognition, and is sufficiently general to be applicable to a wide variety of imaging studies. 
%
Here it is adapted for simultaneous registration and temporal ordering of images from cross-sectional studies of developmental dynamics, with the main objective of revealing stereotypic developmental trajectories.
%
We analyze three data sets from a study of {\it Drosophila} embryogenesis, one of the best experimental models for studies of developmental dynamics \cite{jaeger2012drosophila}.
%
%The first data set, shown in Figure~\ref{subfig:raw_data1}, is relatively simple and will allow us to both illustrate and validate this approach.
%
%In the second data set, shown in Figure~\ref{subfig:raw_data2}, the dynamics are significantly more complex.
%
%Here, we will show that we can recover a trajectory that is qualitatively consistent with current knowledge.
%
%The third set consists of data collected from both wild type and mutant embryos, where joint processing of the data will help to visualize the different sources of variability along distinct developmental paths.

@@

%%%YGK  when I read this I think of Eiffel tower pictures, not of ours - there is a transition... careful where we do it.
%
%Our goal is to organize data sets from cross-sectional studies of developmental processes in a meaningful and informative way.
%
%We will show that dimensionality reduction algorithms can simultaneously register and temporally order the images to give us a meaningful picture of the underlying developmental dynamics.
%
The approach is based on vector diffusion maps \cite{singer2012vector}, a manifold learning
technique developed for data sets which contain two types of sources of variability:
geometric symmetries (such as translations and rotations of the images due to the experimental setup) which one would like to factor out,
and ``additional" directions of variability (such as temporal dynamics) which one wants to uncover.
%
Vector diffusion maps combine two algorithms, {\em angular synchronization} \cite{singer2011angular} for image registration and {\em diffusion maps} \cite{coifman2005geometric} for extracting intrinsic low-dimensional structure in data, into a single computation that allows us to simultaneously register and temporally order our images.
%

Vector diffusion maps has been used to classify images from cryo-electron microscopy (cryo-EM) experiments, where each image is a projection of a randomly oriented molecule.
%
%%%YGK to imply which rotations we need to have discussed it above
%%
Here, we will use the algorithm to register images of {\it Drosophila} embryos with respect to rotations and translations, as well as uncover the main direction of variability {\it after} removing translational and rotational symmetries.
%
We assume that, in these sets of images, the main direction of variability is parameterized by the developmental time of each embryo, so that uncovering this direction will allow us to recover the developmental dynamics.

@@

Both angular synchronization and diffusion maps use pairwise information to uncover global structure in the data.
%
Angular synchronization uses pairwise alignment information to find translations and rotations which are globally consistent, and diffusion maps uses pairwise distances to find parameterizations which are globally informative.
%
Vector diffusion maps combines both steps into a single computation (see \SI).
%
%The algorithm uses information about pairwise alignments {\em and} pairwise distances to find a global parameterization of the data, while registering images which are ``close'' .
%
%It is thus perfectly suited for our task of registering and temporally ordering cross-sectional imaging data.
%

@@

Previous studies established that throughout the third hour of development, the nuclear localization of Dl is unimodal and peaked at the ventral side of the embryo, reviewed in \cite{rushlow2012temporal}.
%
The spatiotemporal pattern of dpERK is more complex: dpERK is first detected in two lateral stripes, which increase in intensity towards the end of the third hour of development \cite{lim2013kinetics}.
%
This bimodal pattern is then augmented by a peak at the dorsal side of the embryo.
%
Both of these features become readily apparent when images are registered and ordered using vector diffusion maps (see \fig~\ref{subfig:ordered_data1}): the peak of nuclear Dl is consistent throughout the images and the intensity of the dpERK signal in the lateral regions of the embryo increases monotonically across the data set.
%

For this data set, the quality of the temporal ordering provided by vector diffusion maps can be assessed by comparing with the ordering obtained through an independent approach.
%
During the third hour of development, the embryo cellularizes: in this process, nuclei located under a common plasma membrane are enclosed by lateral membranes which grow inward, separating the nuclei into individual cells.
%
The highly reproducible kinetics of lateral membrane growth can be used to estimate the age of each embryo to within 1--3 minutes \cite{figard2013plasma} (see \fig~S2).
%
The Spearman rank correlation coefficient between the orderings obtained based on the progress of cellularization and vector diffusion maps is 0.91, indicating that vector diffusion maps can accurately recover the temporal order.
%
Thus, we have validated a dimensionality reduction approach to temporal ordering and registration of data on a data set where image orientation and ordering can be assessed independently.

@@

%
During this window, the ventral furrow is formed, where the ventral side of the embryo buckles towards the center of the embryo, bringing in the future muscle cells.
%
dpERK appears as two lateral peaks at the ventral side of the embryo.
%
These two peaks merge together during invagination, eventually forming (together with Twi) a characteristic ``omega'' shape.
%
During this time window, germband extension causes cells from the ventral side to move towards the posterior pole of the embryo, and then wrap around to the dorsal side.
%
At the end of this process, cells which were originally on the ventral and posterior side of the embryo find themselves on the dorsal side, causing similar ``omega'' shaped patterns to arise on the dorsal side; these patterns are most readily seen in the last image of \fig~\ref{subfig:averaged_data2}.

@@

When time is the main source of variability, we have shown vector diffusion maps can recover developmental progression from cross-sectional data.
%
However, vector diffusion maps can also provide an organization of data which contains multiple sources of variability.
%
To illustrate this, we collected a data set consisting of $42$ embryos undergoing cellularization during the third hour of development (see \fig~S3).
%
Half of the embryos are wild type, and half of the embryos are mutants where dpERK is expressed in one broad peak at the ventral side.
%
Again, each image is an optical cross-section of an embryo fixed at a different developmental time, and the embryos have been stained for nuclei (gray), Dl (green), and dpERK (red).
%
We expect two sources of variability in this data set: differences due to developmental time, and differences between mutant and wild type dynamics.

We computed the vector diffusion maps embedding for our data, and the eigenvalue spectrum (see \fig~S6) indicates that two coordinates may be required for a thorough description of the data. 
%
\fig~\ref{subfig:projection_data3} shows the data projected onto the first two vector diffusion maps coordinates, where each point in this projection corresponds to one image in our data set.
%%%
%%%YGK we need something about dmap eignvalues and how many are not harmonics
%
Two distinct branches appear to emerge in this two-dimensional embedding space.

@@

We presented a unified approach to temporal ordering and registration of images in cross-sectional studies of developmental dynamics. 
%
To the best of our knowledge, algorithmic approaches to these two tasks have been explored largely independently of each other. 
%
In particular, temporal ordering of imaging data sets was done with a significant amount of human supervision and using registered images as a starting point \cite{yuan2014automated, surkova2008characterization}.  
%
In parallel, ordering of large-scale cross-sectional data was done in the context of molecular profiling studies, in which data are vectors describing the amounts of different chemical species such as mRNA abundances \cite{anavy2014blind, trapnell2014dynamics, gupta2008extracting}. 
%
Automated temporal ordering of such data sets was accomplished by first projecting the data on a low-dimensional subspace spanned by the leading principal or independent components, and then solving a traveling salesman problem or constructing a minimum spanning tree on the projected data to order multiple snapshots. 
%
The approach presented here is different because it addresses image registration and ordering in a single step and orders the data without first projecting them on a low-dimensional subspace. 
%
The task of image registration has also been widely studied \cite{zitova2003image}, for applications such as face recognition \cite{rowley1998rotation}, medical image registration \cite{hajnal2010medical}, and texture classification \cite{greenspan1994rotation}.
%
In contrast to most of the existing approaches to image registration which rely on the knowledge about some landmarks in the data \cite{ian1998statistical} (such as the eyes in face recognition applications \cite{zhao2003face}), algorithms based on angular synchronization (such as vector diffusion maps) can register images even in the absence of such information. 
%TODO: look at Eisen Cell paper about registration

The presented approach was demonstrated on three different experimental data sets, collected specifically for this study. 
%
In the first example, correct orientation and temporal order of images was known {\it a priori}, providing a test case for the algorithm in which both of these features were accurately recovered. 
%
For this particular data set, a simpler approach would have sufficed. 
%
Indeed, because of the relative simplicity of developmental dynamics (monotonic increase in the amplitude of a localized signaling pattern and relatively small tissue deformation), temporal order could have been inferred by projection of registered images on the first principal component of the data set. 
%
This approach was clearly not sufficient for the second data set (see \fig~S4), collected from embryos during the time window where both the tissue morphology and signaling patterns are rapidly changing.  
%
However, vector diffusion maps could readily extract a nontrivial spatiotemporal trajectory. 
%
Finally, the algorithm was applied to a heterogeneous data set, composed of images from embryos of different genotypes. 
%
In this case, low-dimensional projection of registered data revealed distinct paths that correlated with the developmental trajectories of the mutant and wild type populations.

@@

\section{Development time during cellularization} 
%
The progression of membrane invagination during cellularization was used to assign time point to each embryo in the third hour of development. Time-lapse imaging of Sqh-GFP embryo was used to generate correlation curve between the membrane ingression length and the developmental time (n=7) (\fig S1B). The rate at which membrane invaginates was highly reproducible from embryo to embryo. In the experiments with fixed embryos, we measured the membrane length by co-staining embryos with a membrane marker (Sqh-GFP) or by taking phase-contrast images (\fig S1A). We used the correlation curve to assign each embryo to a time window of 2--3~minutes.

@@

\bibliographystyle{pnas2}
\bibliography{background_reading/references,../../references/references}

%\begin{thebibliography}{}

%\end{thebibliography}

>>>>

Imaging studies provide unique insights into the dynamics of developmental processes. 
%
A large number of imaging studies are based on a cross-sectional experimental design, where developmental progress is arrested in a population of embryos, each of which is at a different point along its developmental trajectory. 
%
The goal is then to temporally order the data to reconstruct representative developmental dynamics from snapshots provided by the collection of embryos.
%
Images of different biological samples must first be registered before they can be temporally ordered.
%
When such data sets are large, noisy, and/or if the developmental changes are subtle, these tasks can be difficult to perform manually.
%
We present an automated approach to registering and temporally ordering data from such cross-sectional studies.
%
The approach is based on vector diffusion maps, a broadly applicable manifold learning technique that does not require {\it a priori} knowledge of image features or developmental dynamics.
%
%Furthermore, these techniques can separate multiple trajectories within the same data set.
%
We demonstrate the utility of these methods using a collection of images from a cross-sectional study of cell signaling during dorsoventral patterning of the {\it Drosophila} embryo.
\end{abstract}

@@

Image registration has been studied in a variety of contexts \cite{zitova2003image, rowley1998rotation, hajnal2010medical, greenspan1994rotation, zhao2003face}, 
and most registration algorithms are based on landmarks, such as the head and fins in \fig~\ref{fig:fish}.
%
In contrast, our approach requires no such information. 
%
It is based on a manifold learning algorithm \cite{singer2012vector} which simultaneously addresses the problems of registration and temporal ordering. 
%
This algorithm is one of many nonlinear dimensionality reduction techniques that have been developed over the past decade \cite{Belkin2003, coifman2005geometric, coifman2006geometric, tenenbaum2000global, roweis2000nonlinear}. 
%
This technique has been used for data mining in
applications ranging from analysis of cryo-electron microscopy (cryo-EM) images of individual molecules to face recognition, and is sufficiently general to be applicable to a wide variety of biological imaging studies. 
%
Here the algorithm is adapted for simultaneous registration and temporal ordering of images from cross-sectional studies of developmental dynamics, with the main objective of revealing stereotypic developmental trajectories.
%
To illustrate our approach, we analyze three data sets from a study of {\it Drosophila} embryogenesis, one of the best experimental models for studies of developmental dynamics \cite{jaeger2012drosophila}.
%
%The first data set, shown in Figure~\ref{subfig:raw_data1}, is relatively simple and will allow us to both illustrate and validate this approach.
%
%In the second data set, shown in Figure~\ref{subfig:raw_data2}, the dynamics are significantly more complex.
%
%Here, we will show that we can recover a trajectory that is qualitatively consistent with current knowledge.
%
%The third set consists of data collected from both wild type and mutant embryos, where joint processing of the data will help to visualize the different sources of variability along distinct developmental paths.

@@

%%%YGK  when I read this I think of Eiffel tower pictures, not of ours - there is a transition... careful where we do it.
%
%Our goal is to organize data sets from cross-sectional studies of developmental processes in a meaningful and informative way.
%
%We will show that dimensionality reduction algorithms can simultaneously register and temporally order the images to give us a meaningful picture of the underlying developmental dynamics.
%
Our approach is based on vector diffusion maps \cite{singer2012vector}, a manifold learning
technique developed for data sets which contain two types of sources of variability:
geometric symmetries (such as translations and rotations of the images due to the experimental setup) which one would like to factor out,
and ``additional" directions of variability (such as temporal dynamics) which one wants to uncover.
%
Vector diffusion maps combine two algorithms, {\em angular synchronization} \cite{singer2011angular} for image registration and {\em diffusion maps} \cite{coifman2005geometric} for extracting intrinsic low-dimensional structure in data, into a single computation that allows us to simultaneously register and temporally order our images.
%
%Vector diffusion maps has been used to classify images from cryo-electron microscopy (cryo-EM) experiments, where each image is a projection of a randomly oriented molecule.
%
%%%YGK to imply which rotations we need to have discussed it above
%%
We will use the algorithm to register images of {\it Drosophila} embryos with respect to rotations and translations, as well as uncover the main direction of variability {\it after} removing translational and rotational symmetries.
%
We assume that, in these sets of images, the main direction of variability is parameterized by the developmental time of each embryo, so that uncovering this direction will allow us to recover the developmental dynamics.

@@

Both angular synchronization and diffusion maps use pairwise information to uncover global structure in the data.
%
Angular synchronization uses pairwise alignment information to find translations and rotations which are globally consistent, and diffusion maps uses pairwise distances to find parameterizations which are globally informative.
%
Vector diffusion maps combines both steps into a single computation (see \SI), simultaneously addressing the tasks of registration and temporal ordering.
%
%The algorithm uses information about pairwise alignments {\em and} pairwise distances to find a global parameterization of the data, while registering images which are ``close'' .
%
%It is thus perfectly suited for our task of registering and temporally ordering cross-sectional imaging data.
%

@@

Previous studies established that throughout the third hour of development, the nuclear localization of Dl is unimodal and peaked at the ventral side of the embryo (reviewed in \cite{rushlow2012temporal}).
%
The spatiotemporal pattern of dpERK is more complex: dpERK is first detected in two lateral stripes, which increase in intensity towards the end of the third hour of development \cite{lim2013kinetics}.
%
This bimodal pattern is then augmented by a peak at the dorsal side of the embryo.
%
Both of these features become readily apparent when images are registered and ordered using vector diffusion maps (see \fig~\ref{subfig:ordered_data1}): the peak of nuclear Dl is consistent throughout the images and the intensity of the dpERK signal in the lateral regions of the embryo increases monotonically across the data set.
%

For this data set, the quality of the temporal ordering provided by vector diffusion maps can be assessed by comparing with the ordering obtained through an independent approach.
%
During the third hour of development, the embryo cellularizes: in this process, nuclei located under a common plasma membrane are enclosed by lateral membranes which grow inward, separating the nuclei into individual cells.
%
The highly reproducible kinetics of lateral membrane growth can be used to estimate the age of each embryo to within 2--3 minutes \cite{figard2013plasma} (see \fig~S2).
%
The Spearman rank correlation coefficient between the orderings obtained based on the progress of cellularization and vector diffusion maps is 0.91, indicating that vector diffusion maps can accurately recover the temporal order.
%
Thus, we have validated a dimensionality reduction approach to temporal ordering and registration on a data set where image orientation and ordering can be assessed independently.

@@

%
During this window, the ventral furrow is formed, where the ventral side of the embryo buckles towards the center of the embryo, bringing in the future muscle cells.
%
dpERK appears as two lateral peaks at the ventral side of the embryo.
%
These two peaks merge together during invagination, eventually forming (together with Twi) a characteristic ``omega'' shape.
%
During this time window, germband extension causes cells from the ventral side to move towards the posterior pole of the embryo, and then wrap around to the dorsal side \cite{leptin2005gastrulation}.
%
At the end of this process, cells which were originally on the ventral and posterior side of the embryo find themselves on the dorsal side, causing similar ``omega'' shaped patterns to arise on the dorsal side; these patterns are most readily seen in the last image of \fig~\ref{subfig:averaged_data2}.

@@

When time is the main source of variability, we have shown vector diffusion maps can recover developmental progression from cross-sectional data.
%
However, vector diffusion maps can also provide an organization of data which contains multiple sources of variability.
%
To illustrate this point, we collected a data set consisting of $42$ embryos undergoing cellularization during the third hour of development (see \fig~S3).
%
Half of the embryos are wild type, and half of the embryos are mutants where dpERK is expressed in one broad peak at the ventral side.
%
Again, each image is an optical cross-section of an embryo fixed at a different developmental time, and the embryos have been stained for nuclei (gray), Dl (green), and dpERK (red).
%
We expect two sources of variability in this data set: differences due to developmental time, and differences between mutant and wild type dynamics.

We computed the vector diffusion maps embedding for our data, and the eigenvalue distribution (see \fig~S6) indicates that two coordinates may be required for a thorough description of the data. 
%
\fig~\ref{subfig:projection_data3} shows the data projected onto the first two vector diffusion maps coordinates, where each point in this projection corresponds to one image in our data set.
%%%
%%%YGK we need something about dmap eignvalues and how many are not harmonics
%
Two distinct branches appear to emerge in this two-dimensional embedding space.

@@

We presented a unified approach to temporal ordering and registration of images in cross-sectional studies of developmental dynamics. 
%
To the best of our knowledge, algorithmic approaches to these two tasks have been explored largely independently of each other. 
%
In particular, temporal ordering of imaging data sets was done with a significant amount of human supervision and using registered images as a starting point \cite{yuan2014automated, surkova2008characterization}.  
%
In parallel, ordering of large-scale cross-sectional data was done in the context of molecular profiling studies, in which data are vectors describing the amounts of different chemical species, such as mRNA abundances \cite{anavy2014blind, trapnell2014dynamics, gupta2008extracting}. 
%
Automated temporal ordering of such data sets was accomplished by first projecting the data on a low-dimensional subspace spanned by the leading principal or independent components, and then solving a traveling salesman problem or constructing a minimum spanning tree on the projected data to order multiple snapshots. 
%
The approach presented here is different because it addresses image registration and ordering in a single step and orders the data without first projecting them on a low-dimensional subspace. 
%
The task of image registration has also been widely studied \cite{zitova2003image}, for applications such as face recognition \cite{rowley1998rotation}, medical image registration \cite{hajnal2010medical}, and texture classification \cite{greenspan1994rotation}.
%
In contrast to most of the existing approaches to image registration which rely on the knowledge about some landmarks in the data \cite{ian1998statistical} (such as the eyes in face recognition applications \cite{zhao2003face}), algorithms based on angular synchronization (such as vector diffusion maps) can register images even in the absence of such information. 
%TODO: look at Eisen Cell paper about registration

The presented approach was demonstrated on three different experimental data sets, collected specifically for this study. 
%
In the first example, correct orientation and temporal order of images was known {\it a priori}, providing a test case for the algorithm in which both of these features were accurately recovered. 
%
For this particular data set, a simpler approach would have sufficed. 
%
Indeed, because of the relative simplicity of developmental dynamics (monotonic increase in the amplitude of a localized signaling pattern and relatively small tissue deformation), temporal order could have been inferred by projection of registered images on the first principal component of the data set (see \fig~S7). 
%
This approach was clearly not sufficient for the second data set (see \fig~S8), collected from embryos during the time window where both the tissue morphology and signaling patterns are rapidly changing.  
%
However, vector diffusion maps could readily extract a nontrivial spatiotemporal trajectory. 
%
Finally, the algorithm was applied to a heterogeneous data set, composed of images from embryos of different genotypes. 
%
In this case, low-dimensional projection of registered data revealed distinct paths that correlated with the developmental trajectories of the mutant and wild type populations.

@@

\section{Development time during cellularization} 
%
The progression of membrane invagination during cellularization was used to assign time point to each embryo in the third hour of development. Time-lapse imaging of Sqh-GFP embryo was used to generate correlation curve between the membrane ingression length and the developmental time (n=7) (\fig S2B). The rate at which membrane invaginates was highly reproducible from embryo to embryo. In the experiments with fixed embryos, we measured the membrane length by co-staining embryos with a membrane marker (Sqh-GFP) or by taking phase-contrast images (\fig S2A). We used the correlation curve to assign each embryo to a time window of 2--3~minutes.

@@

%\bibliographystyle{pnas2}
%\bibliography{background_reading/references,../../references/references}

\begin{thebibliography}{10}

\bibitem{diggle2002analysis}
Diggle P, Heagerty P, Liang K~Y, Zeger S (2002) \textit{Analysis of
  longitudinal data} (Oxford University Press).

\bibitem{roelens2013live}
Roelens B, Chanes J~D~L~H, Dostatni N (2013) Live imaging of bicoid-dependent
  transcription in \textit{{D}rosophila} embryos. \textit{Curr Biol}
  23:1--5.

\bibitem{keller2013imaging}
Keller P~J (2013) Imaging morphogenesis: technological advances and biological
  insights. \textit{Science} 340(6137).

\bibitem{jaeger2004dynamic}
Jaeger J, Surkova S, Blagov M, Janssens H, Kosman D, \textit{et~al.} (2004)
  Dynamic control of positional information in the early \textit{{D}rosophila}
  embryo. \textit{Nature} 430(6997):368--371.

\bibitem{fowlkes2008quantitative}
Fowlkes C~C, Hendriks C~L~L, Ker{\"a}nen S~V, Weber G~H, R{\"u}bel O,
  \textit{et~al.} (2008) A quantitative spatiotemporal atlas of gene expression
  in the \textit{{D}rosophila} blastoderm. \textit{Cell} 133(2):364--374.

\bibitem{ng2012large}
Ng L~L, Sunkin S~M, Feng D, Lau C, Dang C, \textit{et~al.} (2012) Large-scale
  neuroinformatics for in situ hybridization data in the mouse brain.
  \textit{Int Rev Neurobiol} 104:159--182.

\bibitem{richardson2014emage}
Richardson L, Stevenson P, Venkataraman S, Yang Y, Burton N, \textit{et~al.}
  (2014) Emage: Electronic mouse atlas of gene expression, \textit{Mouse
  Molecular Embryology} (Springer), pp. 61--79.

\bibitem{castro2009automatic}
Castro C, Luengo-Oroz M, Desnoulez S, Duloquin L, Fernandez-de Manuel L,
  \textit{et~al.} (2009) An automatic quantification and registration strategy
  to create a gene expression atlas of zebrafish embryogenesis,
  \textit{Conf Proc IEEE Eng Med Biol Soc}, pp. 1469--1472.

\bibitem{zitova2003image}
Zitova B, Flusser J (2003) Image registration methods: a survey. \textit{Image Vis Comput} 21(11):977--1000.

\bibitem{rowley1998rotation}
Rowley H~A, Baluja S, Kanade T (1998) Rotation invariant neural network-based
  face detection, \textit{Computer Vision and Pattern Recognition, 1998.
  Proceedings. 1998 IEEE Computer Society Conference on} (IEEE), pp. 38--44.

\bibitem{hajnal2010medical}
Hajnal J~V, Hill D~L (2010) \textit{Medical image registration} (CRC press).

\bibitem{greenspan1994rotation}
Greenspan H, Belongie S, Goodman R, Perona P (1994) Rotation invariant texture
  recognition using a steerable pyramid, \textit{Pattern Recognition, 1994.
  Vol. 2-Conference B: Computer Vision \& Image Processing., Proceedings of the
  12th IAPR International. Conference on} (IEEE), volume~2, pp. 162--167.

\bibitem{zhao2003face}
Zhao W, Chellappa R, Phillips P~J, Rosenfeld A (2003) Face recognition: A
  literature survey. \textit{ACM Computing Surveys (CSUR)} 35(4):399--458.

\bibitem{singer2012vector}
Singer A, Wu H~T (2012) Vector diffusion maps and the connection laplacian.
  \textit{Commun Pure Appl Math} 65(8):1067--1144.

\bibitem{Belkin2003}
Belkin M, Niyogi P (2003) Laplacian eigenmaps for dimensionality reduction and
  data representation. \textit{Neural Comput} 15:1373--1396.

\bibitem{coifman2005geometric}
Coifman R~R, Lafon S, Lee A~B, Maggioni M, Nadler B, \textit{et~al.} (2005)
  Geometric diffusions as a tool for harmonic analysis and structure definition
  of data: Diffusion maps. \textit{Proc Natl Acad Sci USA} 102(21):7426--7431.

\bibitem{coifman2006geometric}
Coifman R~R, Lafon S (2006) Geometric harmonics: a novel tool for multiscale
  out-of-sample extension of empirical functions. \textit{Appl Comput Harmon Anal} 21(1):31--52.

\bibitem{tenenbaum2000global}
Tenenbaum J~B, De~Silva V, Langford J~C (2000) A global geometric framework for
  nonlinear dimensionality reduction. \textit{Science} 290(5500):2319--2323.

\bibitem{roweis2000nonlinear}
Roweis S~T, Saul L~K (2000) Nonlinear dimensionality reduction by locally
  linear embedding. \textit{Science} 290(5500):2323--2326.

\bibitem{jaeger2012drosophila}
Jaeger J, Manu, Reinitz J (2012) \textit{{D}rosophila} blastoderm patterning.
  \textit{Curr Opin Genet Dev} 22(6):533 -- 541, genetics of system biology.

\bibitem{singer2011angular}
Singer A (2011) Angular synchronization by eigenvectors and semidefinite
  programming. \textit{Appl Comput Harmon Anal}
  30(1):20--36.

\bibitem{ahuja2007template}
Ahuja S, Kevrekidis I~G, Rowley C~W (2007) Template-based stabilization of
  relative equilibria in systems with continuous symmetry. \textit{Journal of
  Nonlinear Science} 17(2):109--143.

\bibitem{ian1998statistical}
Dryden I~L, Mardia K (1998) \textit{Statistical shape analysis}, Wiley series
  in probability and statistics: Probability and statistics (J. Wiley).

\bibitem{foe1983studies}
Foe V~E, Alberts B~M (1983) Studies of nuclear and cytoplasmic behaviour during
  the five mitotic cycles that precede gastrulation in \textit{{D}rosophila}
  embryogenesis. \textit{J Cell Sci} 61(1):31--70.

\bibitem{rushlow2012temporal}
Rushlow C~A, Shvartsman S~Y (2012) Temporal dynamics, spatial range, and
  transcriptional interpretation of the dorsal morphogen gradient.
  \textit{Curr Opin Genet Dev} 22(6):542--546.

\bibitem{lim2013kinetics}
Lim B, Samper N, Lu H, Rushlow C, Jim{\'e}nez G, \textit{et~al.} (2013)
  Kinetics of gene derepression by erk signaling. \textit{Proc Natl Acad Sci USA} 110(25):10330--10335.

\bibitem{figard2013plasma}
Figard L, Xu H, Garcia H~G, Golding I, Sokac A~M (2013) The plasma membrane
  flattens out to fuel cell-surface growth during \textit{{D}rosophila}
  cellularization. \textit{Dev Cell} 27(6):648--655.

\bibitem{shlens2005tutorial}
Shlens J (2005) A tutorial on principal component analysis. \textit{Systems
  Neurobiology Laboratory, University of California at San Diego} .

\bibitem{leptin2005gastrulation}
Leptin M (2005) Gastrulation movements: the logic and the nuts and bolts.
  \textit{Dev Cell} 8(3):305--320.

\bibitem{yuan2014automated}
Yuan L, Pan C, Ji S, McCutchan M, Zhou Z~H, \textit{et~al.} (2014) Automated
  annotation of developmental stages of \textit{{D}rosophila} embryos in images
  containing spatial patterns of expression. \textit{Bioinformatics}
  30(2):266--273.

\bibitem{surkova2008characterization}
Surkova S, Kosman D, Kozlov K, Myasnikova E, Samsonova A~A, \textit{et~al.}
  (2008) Characterization of the \textit{{D}rosophila} segment determination
  morphome. \textit{Dev Biol} 313(2):844--862.

\bibitem{anavy2014blind}
Anavy L, Levin M, Khair S, Nakanishi N, Fernandez-Valverde S~L, \textit{et~al.}
  (2014) Blind ordering of large-scale transcriptomic developmental
  timecourses. \textit{Development} 141(5):1161--1166.

\bibitem{trapnell2014dynamics}
Trapnell C, Cacchiarelli D, Grimsby J, Pokharel P, Li S, \textit{et~al.} (2014)
  The dynamics and regulators of cell fate decisions are revealed by
  pseudotemporal ordering of single cells. \textit{Nat Biotechnol} 32(4):381--386.

\bibitem{gupta2008extracting}
Gupta A, Bar-Joseph Z (2008) Extracting dynamics from static cancer expression
  data. \textit{IEEE/ACM Trans Comput Biol Bioinform} 5(2):172--182.

\bibitem{singer2011three}
Singer A, Shkolnisky Y (2011) Three-dimensional structure determination from
  common lines in cryo-em by eigenvectors and semidefinite programming.
  \textit{SIAM J Imaging Sci} 4(2):543--572.

\bibitem{silverman2005functional}
Ramsay J~O, Silverman B~W (2005) \textit{Functional Data Analysis}, Springer
  Series in Statistics (Springer).

\bibitem{krzic2012multiview}
Krzic U, Gunther S, Saunders T~E, Streichan S~J, Hufnagel L (2012) Multiview
  light-sheet microscope for rapid in toto imaging. \textit{Nat Methods}
  9(7):730--733.

\bibitem{ichikawa2014live}
Ichikawa T, Nakazato K, Keller P~J, Kajiura-Kobayashi H, Stelzer E~H,
  \textit{et~al.} (2014) Live imaging and quantitative analysis of gastrulation
  in mouse embryos using light-sheet microscopy and 3d tracking tools.
  \textit{Nat Protoc} 9(3):575--585.

\bibitem{rubel2010coupling}
R{\"u}bel O, Ahern S, Bethel E, Biggin M~D, Childs H, \textit{et~al.} (2010)
  Coupling visualization and data analysis for knowledge discovery from
  multi-dimensional scientific data. \textit{Procedia Comput Sci}
  1(1):1757--1764.

\end{thebibliography}
